{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}

### 2017-01-06 proposal/410.tex
410 resultce

This chapter proposes my plan to evaluate Landslide's effectiveness as a debugging aid for students in an educational setting.
This is the second of the three projects I am proposing for this thesis, and is currently ongoing work.

@@

>>>>

This chapter proposes my plan to evaluate Landslide's effectiveness as a debugging aid for students in an educational setting.
This is the second of the three projects I am proposing for this thesis, and is currently ongoing work.
What remains to be done is detailed in Section~\ref{sec:grading}.

@@

In the Spring 2015, Fall 2015, and Spring 2016 semesters,
I've made Landslide (with Quicksand) available to 15-410 students during the last week of P2.
I introduced stateless model checking in a guest lecture given to the class,
and required students to pass the thread library hurdle \cite{thrlib}
to ensure a minimum of basic functionality necessary to use Landslide.

To analyze its effectiveness as a debugging tool for students,
I configured it to record a snapshot of the results of each use by the students:
which options were used (test program and time limit),
a snapshot of the student’s code, and the result of the test
(completed, timed out, bug found, or ctrl-C’ed).
I manually interpreted the snapshots to determine:
\begin{itemize}
	\item How many unique bugs the student found (discounting multiple runs that produced ``the same'' bug);
	\item How many of those bugs were deterministic versus concurrency bugs (did Landslide find the bug on the first interleaving or did it need to test multiple);
	\item How many of each category of bug the students fixed (determined when a subsequent run of Landslide on the test case failed to find the bug again).
\end{itemize}

{\bf Results -- bugs found.}
Of 90 two-student groups in those 3 semesters, 47 tested their thread libraries with Landslide.
Table~\ref{tab:this-table-sucks-but-it's-the-best-i-got} shows
that Landslide found in total 44 deterministic bugs and 85 concurrency bugs.
It found at least one concurrency bug for 32 groups (68\%),
24 of whom (51\%) were able to fix at least one, verifying their update
with a succesful re-run of the same test.

The table's data presentation is somewhat confusing:
the dependent columns count the number of {\em groups}, categorized by how many bugs they found, not the raw number of {\em bugs},
which I instead indicate in the ``Total bugs'' row at the bottom.
So, the number of total bugs is equal to the sum of the products between each cell in that column and its corresponding number of bugs,
e.g., $44 = 27\times0 + 6\times1 + 9\times2 + 2\times3 + 1\times4 + 2\times5$.
Sorry about that; I'll try to find a better way to present this data in the thesis.

\begin{table}[h]
	\begin{center}
	\begin{tabular}{r|cc|cc}
	%det bugs found det bugs fixed  races fixed     races found
	& \multicolumn{4}{c}{Number of groups} \\
	Number & \multicolumn{2}{c|}{Deterministic bugs} & \multicolumn{2}{c}{Concurrency bugs} \\
	of bugs & found & fixed & found & fixed \\
	\hline
	0       & 27    & 32    & 15    & 23    \\
	1       & 6     & 5     & 7     & 11    \\
	2       & 9     & 6     & 12    & 8     \\
	3       & 2     & 1     & 7     & 2     \\
	4       & 1     & 1     & 3     & 1     \\
	5       & 2     & 2     & 2     & 1     \\
	11      &       &       & 1     & 1     \\
	\hline
	%Total groups
	%       & 47    & 47    & 47    & 47    \\
	Total bugs
	        & 44    & 34    & 85    & 53    \\
	\end{tabular}
	\end{center}
	\caption{Summary of how many bugs were found and/or fixed by how many groups.
	Each column counts how many groups found/fixed the number of bugs in each row;
	for instance, two groups found 5 concurrency bugs, one of which fixed all 5.}
	\label{tab:this-table-sucks-but-it's-the-best-i-got}
\end{table}

{\bf Results -- impact on grades.}
Next, I investigated the impact Landslide ultimately had on students' performance.
I collected the ultimate P2 and P3 grades from the last 6 semesters, forming three categories:
students from the past 3 semesters who volunteered to use Landslide,
students from those semesters who opted out,
and students from the other 3 semesters, for whom Landslide was not available at all.
I hoped to measure a correlation between use of Landslide and ultimate P2 grades,
which would show that it helps students submit more correct P2s,
and/or a correlation with ultimate P3 grades,
which would (tentatively) show that it teaches better ways of thinking that students retain for future projects (i.e., measuring actual learning).

Unfortunately, this turned out to be a negative result, as shown in Figure~\ref{fig:grade-distribution}.
The distribution of Landslide users' ultimate grades is indistinguishable from that of students for whom Landslide was unavailable.
While there was a 3\% increase in P2 grades between opt-out-ers and Landslide users, it's prone to selection bias:
perhaps those who opted out of Landslide were already the worst of the class who simply didn't have any free time for it.
The distribution of P3 grades is indistinguishable as well.
In the thesis, I will discuss possible factors contributing to this failure, and how they might be controlled for in future work,
although I am not planning to run any further experiments in the limited time I have.

\begin{figure}[h]
	\begin{center}
	\includegraphics[width=0.96\textwidth]{p2-p3-distribution.pdf}
	\end{center}
	\caption{Distribution of P2 (thrlib) and P3 (kernel) grades among students who did or didn't use Landslide.}
	\label{fig:grade-distribution}
\end{figure}

{\bf Future work.}
One evaluation question remains which I will answer in the thesis:
what is the accuracy of Landslide's automation heuristics (see previous section)?
In other words, were false-positive infinite loops or deadlocks ever reported
where Landslide failed to recognize an ad-hoc synchronization?
(I expect a near 100\% success rate,
having tuned the heuristics by hand on an (independent) set of P2 submissions already.)
